Software Clusterings with Vector Semantics and the Call Graph

### 2019-06-25 manuscript/manuscript.tex
Adjustments according to reviewers' comments

@@

\begin{table*}
  \caption{Experimental Results for Linux 4.21}
    \label{tab:evaluation}
    \begin{tabular}{lrrrrrrr}
    \toprule
    Algorithm & Feature Dimension & \# Clusters & Size Range & $\bar x$ & $\sigma$ & Median & MoJo Distance \\
    \midrule
    \textsc{acdc} \cite{acdc} & -- & 9055 & 1 -- 4245 & 5 & 46 & 2 & 33694\\
    Average Linkage \cite{average} & \emph{300} & \emph{21} & 1--3406 & 163 & 725 & 1 & 2092 \\
    Complete Linkage \cite{complete} & \emph{300} & \emph{21} & 1--1529 & 163 & 412 & 19 & 1710 \\
    \textsc{limbo} \cite{limbo} & 12317 &\emph{21} & 50--1810 & 163 & 375 & 50  & 1482 \\

@@

\paragraph{Embeddings}

For the prepossessing stage we use the \textsc{nlp} package spaCy \cite{spacy} which provided us with very useful tools for preprocessing. 
After prepossessing, we use Gensim \cite{gensim} to train the Doc2Vec model. 
First of all, we tokenize the code and remove the stop-words. 
After that, for each token, we split it into its constituent parts using dynamic programming \cite{wordninja} and lemmatize 
each individual sub-token. 
For example, the method \texttt{\_\_zone\_seqlock\_init} corresponds to \texttt{zone}, \texttt{seqlock} and \texttt{init}, 
\texttt{inprogress} is split into \texttt{in} and \texttt{progress} and \texttt{literals} becomes \texttt{literal}.

\paragraph{Call Graph Generation} 
The codebase is processed by a Static Graph Analyzer. 
In our approach, focusing on C projects, we use \emph{CScout} \cite{cscout} in order to generate the directed call graph through function calls between the files. 
Each source code file is assigned to a module. 
Modules can be user defined or automatically generated by their respective directories at a desired directory tree depth.

\paragraph{Combining semantics and the Call Graph} 
The input to the clustering procedure
is the call graph with cosine similarities of the modules as edge weights.
Then we run the Louvain Clustering Algorithm \cite{louvain} in
order to obtain the software clustering via maximizing the modularity function with a greedy approach. 
In case we want to consider edge directionality, we do a bipartite transformation, detect communities and
merge the results with a disjoint-set data structure as stated in reference \cite{malliaros}.

\paragraph{Implementation}  
Our method is implemented in the Python Language using spaCy\cite{spacy}, Gensim \cite{gensim} and NetworkX \cite{nx} and is named \textsc{sade}. The source code and data are available in \cite{source_code} and \cite{call_graph}.

@@

\paragraph{Experimental Setting} Our method was tested on Linux 4.21,
consisting of 20.3 million \textsc{sloc} against Average-Linkage \cite{average}, Complete-Linkage \cite{complete} 
and Ward-Linkage \cite{ward} using the same document embeddings
as well as \textsc{acdc} with structural information \cite{acdc} and \textsc{limbo} \cite{limbo} with $B = 100$, $S = \infty$ and Bag-of-Words features. The results are presented in Table \ref{tab:evaluation} and user-defined parameters are formatted in italics. 
As ground truth, we have used the first level directories as a target clustering and as input we have considered 
the modules of the one-top directories. 
For example, the source code file \texttt{drivers/net/ieee802154\-/mcr20a.c} has a ground truth value of \texttt{drivers} 
and it is considered under the same module as every \texttt{.c} and \texttt{.h} file under \texttt{drivers/net/ieee802154}. 
The MoJo distance measure which counts the minimum number of operations to transform one clustering to another.  
Since Louvain Community Detection produces  different (but very similar) results every time it runs, we ran the simulations 
multiple times and averaged the results. 
The experimentation with a large and complex system may constitute a threat to our findings' validity since the results are not general.

@@

%Their suggestion in \cite{large_study} is a \emph{term frequency--inverse 
% document frequency} (TF-IDF) model serving as a similarity measure between source code 
% entities. 
Work in Software Clustering Algorithms (\textsc{sca}) has been extensively done by many authors in
references \cite{mojo}, \cite{stability}, \cite{acdc} and \cite{limbo} with the \textsc{acdc}
and \textsc{limbo} clustering algorithms. 
In particular, the \textsc{acdc} algorithm leans toward to software components comprehension 
based on subsystem patterns. Their approach considers an initial structure of the system, 
without taking into account semantics, and tries to build comprehensive clusterings of the given ground structure. 
The authors also conducted a study on an older version of  Linux. 
Moreover, Andritsos and Tzerpos in reference \cite{limbo} present an Information-Theoretic 
approach of \textsc{sca} by developing \textsc{limbo} which clusters modules upon inserting their Distributional Cluster Features 
to a B+-tree variant and then applying the \emph{Agglomerative Information Bottleneck} algorithm, merging the 
clusters with the minimum mutual information loss together.

@@

This paper is set out to show how we can use a combination of vector semantics
and information from the call graph in order to produce meaningful clusterings of a software system.
We conclude that we have outperformed state-of-the-art software clusterings and conventional
agglomerative clustering algorithms 
in terms of \emph{extremity}, \emph{authoritativeness} and \emph{stability} 
without even knowing the number of clusters of the ground truth. This evidence 
supports the further usage of vector semantics and the call graph for architecture recovery.
Finally, as future work we mention the use of this project 
with more static analyzers, as well as its use to recover layers in layered
architectural schemata.

>>>>

Work in Software Clustering Algorithms (\textsc{sca}) has been extensively done by many authors in
references \cite{mojo}, \cite{stability}, \cite{acdc} and \cite{limbo} with the \textsc{acdc}
and \textsc{limbo} clustering algorithms. 
In particular, the \textsc{acdc} algorithm leans toward to software components comprehension 
based on subsystem patterns. Their approach considers an initial structure of the system, 
without taking into account semantics, and tries to build comprehensive clusterings of the given ground structure. 
The authors also conducted a study on an older version of  Linux. 
Moreover, Andritsos and Tzerpos in reference \cite{limbo} present an Information-Theoretic 
approach of \textsc{sca} by developing \textsc{limbo} which clusters modules upon inserting their Distributional Cluster Features 
to a B+-tree variant and then applying the \emph{Agglomerative Information Bottleneck} algorithm.

@@

\begin{table*}
  \caption{Experimental Results for Linux 4.21}
  \small
    \label{tab:evaluation}
    \begin{tabular}{lrrrrrrr}
    \toprule
    Algorithm & Feature Dimension & \# Clusters & Size Range & $\bar x$ & $\sigma$ & Median & MoJo Distance \\
    \midrule
    \textsc{acdc} \cite{acdc} & -- & 9055 & 1 -- 4245 & 5 & 46 & 2 & 33694\\
    Average Linkage \cite{average} & \emph{300} & \emph{21} & 1--3406 & 163 & 725 & 1 & 2092 \\
    Complete Linkage \cite{complete} & \emph{300} & \emph{21} & 1--1529 & 163 & 412 & 19 & 1710 \\
    \textsc{limbo} ($B=100, \; S = \infty$) \cite{limbo} & 12317 &\emph{21} & 50--1810 & 163 & 375 & 50  & 1482 \\

@@

\paragraph{Method Overview} The codebase is initially preprocessed and the resulting data is fed into a Skip-Gram model for the initial components, which correspond to nodes in the call graph. Next, we add weights between the nodes with the normalized similarity between them. Finally, we run the Louvain algorithm to obtain a software clustering.

First of all, we tokenize the code and remove the stop-words. 
After that, for each token, we split it into its constituent parts using dynamic programming \cite{wordninja} and lemmatize 
each individual sub-token via using the \textsc{nlp} package spaCy \cite{spacy}. 
For example, the method \texttt{\_\_zone\_seqlock\_init} corresponds to \texttt{zone}, \texttt{seqlock} and \texttt{init}, 
\texttt{inprogress} is split into \texttt{in} and \texttt{progress} and \texttt{literals} becomes \texttt{literal}.
We, then, train a Doc2Vec model using Gensim \cite{gensim}. 
The codebase is also processed by a Static Graph Analyzer. 
In our approach, focusing on C projects, we use \emph{CScout} \cite{cscout} in order to generate the directed call graph through function calls between the files. 
Each source code file is assigned to a module. 
Modules can be user defined or automatically generated by their respective directories at a desired directory tree depth. 
The input to the clustering procedure
is the call graph with cosine similarities of the modules as edge weights.
Then we run the Louvain Clustering Algorithm \cite{louvain} in
order to obtain the software clustering via maximizing the modularity function with a greedy approach. 
In case we want to consider edge directionality, we do a bipartite transformation, detect communities and
merge the results with a disjoint-set data structure as stated in reference \cite{malliaros}.

Our method is implemented in the Python Language using spaCy\cite{spacy}, Gensim \cite{gensim} and NetworkX \cite{nx} and is named \textsc{sade}. The source code and data are available in \cite{source_code} and \cite{call_graph}.

@@

\paragraph{Experimental Setting} Our method was tested on Linux 4.21,
consisting of 20.3 million \textsc{sloc} against Average-Linkage \cite{average}, Complete-Linkage \cite{complete} 
and Ward-Linkage \cite{ward} using the same document embeddings
as well as \textsc{acdc} with structural information \cite{acdc} and \textsc{limbo} \cite{limbo} with Bag-of-Words features. The results are presented in Table \ref{tab:evaluation} and user-defined parameters are formatted in italics. 
As ground truth, we have used the first level directories as a target clustering and as input we have considered 
the modules of the one-top directories. 
For example, the source code file \texttt{drivers/net/ieee802154\-/mcr20a.c} has a ground truth value of \texttt{drivers} 
and it is considered under the same module as every \texttt{.c} and \texttt{.h} file under \texttt{drivers/net/ieee802154}. 
The MoJo distance measure which counts the minimum number of operations to transform one clustering to another.  
Since Louvain Community Detection produces  different (but very similar) results every time it runs, we ran the simulations 
multiple times and averaged the results. 
The experimentation with a large and complex system may constitute a threat to our findings' validity since the results are not general.

@@

@@

This paper is set out to show how we can use a combination of vector semantics
and information from the call graph in order to produce meaningful clusterings of a software system.
We have outperformed state-of-the-art software clusterings and conventional
agglomerative clustering algorithms 
in terms of \emph{extremity}, \emph{authoritativeness} and \emph{stability} 
without even knowing the number of clusters of the ground truth. This evidence 
supports the further usage of vector semantics and the call graph for architecture recovery.
Finally, the further development of evaluation policies with users should be taken into account.
